\documentclass{article}
\usepackage{amsthm, amssymb, amsmath,verbatim}
\usepackage[margin=1in]{geometry}
\usepackage{enumerate}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\N}{\mathbb{N}}



\newtheorem*{claim}{Claim}
\newtheorem{ques}{Question}


\title{CSE 101 Homework 5}
\date{Winter 2023}

\begin{document}

\maketitle

This homework is due on gradescope Friday March 3rd at 11:59pm on gradescope. Remember to justify your work even if the problem does not explicitly say so. Writing your solutions in \LaTeX is recommend though not required.

\begin{ques}[Gameshow Again, 30 points]
Dirk's gameshow from Homework 4 decides to change their rules again. Now Dirk can attempt at most $k$ challenges but \emph{cannot} attempt the same challenge more than once. Give an $O(n\log(n)+kn)$ algorithm to find the strategy that optimizes Dirk's expected winnings. 

Hint: Based on the solution to part (a) from the previous problem you can note that whatever challenges Dirk attempts to try, he should always attempt them in decreasing order of $p_i R_i / (1-p_i)$ (with the order not mattering if there are ties). You can assume this without proof.
\end{ques}

\begin{ques}[Interval Cover Redux, 30 points]
Consider the interval cover from before, but with weights assigned to each interval. You are given a collection $C$ of $n$ intervals in the real line, each with an associated positive weight. Your goal is to find a set $S$ of intervals from $C$ so that no two elements of $S$ overlap and so that subject to this, the sum of the weights of these intervals is as large as possible. Give an $O(n \log(n))$ time algorithm for this problem.
\end{ques}

\begin{ques}[Pitstop Planning, 40 points]
Jake is a racecar driver. In the current race, he needs to take $n$ laps around the course. Unfortunately, his car's performance slowly gets worse each lap he performs without taking a pitstop. In particular, a lap will take time $T_\ell$ if Jake has gone $\ell$ laps since his last one. Unfortunately, Jake's pit crew has only materials to let him perform at most $k$ pitstops over the course of the race.

Give an $O(kn^2)$ time algorithm to determine which laps Jake should take pitstops in so as to minimize his total time to complete all $n$ laps.
\end{ques}

\begin{ques}[Extra credit, 1 point]
Where have I been getting the character names for assignments this quarter from?
\end{ques}

\end{document} 